A thermodynamic view of dusty protoplanetary disks: axisymmetric stability

### 2016-08-28 model.tex
additional discussion in dust edge instability

We can neglect density gradients in Eq. \ref{spm} 
if $r\p_r\ln{\rhog}\sim O(1)$ and $H_\epsilon/\Hgas\gg
\sqrt{\epsilon}\hgas$. For example, if $\epsilon\simeq 0.01$ and
$\hgas\simeq 0.05$, then we require $H_\epsilon/\Hg\gg
5\times10^{-3}$. (So the dust layer thickness $\Hd\gg 5
\times10^{-3}\Hgas$ as well.) Then 
instability requires 
\begin{align}
  \left|\frac{\p\ln{\epsilon}}{\p r}\right| \gtrsim
  \frac{1}{\Htilde}\frac{\left(1+\epsilon\right)}{\sqrt{\epsilon}}. 
\end{align}
That is, if the dust-to-gas ratio varies on a radial lengthscale less
than $O(\Htilde)$, i.e. its vertical lengthscale,  
%$O(\widetilde{H}\sqrt{\epsilon})$ for $\epsilon\ll 1$, or less than
%$O(\widetilde{H}/\sqrt{\epsilon})$ for $\epsilon\gg 1$, 
then the system is potentially unstable.

>>>>

We can neglect density gradients in Eq. \ref{spm} 
if $r\p_r\ln{\rhog}\sim O(1)$ and $H_\epsilon/\Hgas\gg
\sqrt{\epsilon}\hgas$. For example, if $\epsilon\simeq 0.01$ and
$\hgas\simeq 0.05$, then we require $H_\epsilon/\Hg\gg
5\times10^{-3}$. (So the dust layer thickness $\Hd\gg 5
\times10^{-3}\Hgas$ as well.) Then 
instability requires 
\begin{align}\label{dust_edge}
  \left|\frac{\p\ln{\epsilon}}{\p r}\right| \gtrsim
  \frac{1}{\Htilde}\frac{\left(1+\epsilon\right)}{\sqrt{\epsilon}}.
%\simeq
 % \frac{1}{\sqrt{\epsilon}H_\epsilon}, 
\end{align}
%where the second equality applies to $\epsilon\ll 1$. 
That is, if the dust-to-gas ratio varies on a radial lengthscale much less
than $O(\Htilde)$, i.e. its vertical lengthscale,  
%$O(\widetilde{H}\sqrt{\epsilon})$ for $\epsilon\ll 1$, or less than
%$O(\widetilde{H}/\sqrt{\epsilon})$ for $\epsilon\gg 1$, 
then the system is potentially unstable.  
%For $\epsilon\ll 1$ the
%dust-to-gas ratio should vary on a lengthscale shorter than
%$\sqrt{\epsilon}H_\epsilon$. We apply this below.

PPDs typically have $\epsilon$ of $O(10^{-2})$.  Then 
for thin dust layers, $H_\epsilon \simeq \Hd\ll \Hg$, instability
requires $|\p_r\ln{\epsilon}|^{-1}\ll O(0.1\Hg)$, i.e. a very short
radial length-scale in the dust-to-gas ratio. This is unlikely to
occur. On the other hand, if dust is well-mixed, say $\Hd\simeq
0.99\Hg$, then instability only requires
$|\p_r\ln{\epsilon}|^{-1}\lesssim O(10\Hg)$, which is more realistic.
